\section{Content}

\subsection{Useful definitions}

\begin{definition}[Synchronous communication model]\label{def:synch}
An execution $\execution=(w,\source) \in \executionsofmodel{\synchmodel}$
if for all send event $s\in\sendeventsof{e}$, $s+1$ is a receive event
of $e$ and $\source(s+1)=s$.
\end{definition}

\begin{definition}[$\acommunicationmodel$-linearisable MSC]\label{def:linearisable-msc}
An MSC $\mmsc$ is \textit{linearisable} in a communication model $\acommunicationmodel$ if  
$\linearisationsof{\mmsc}{\acommunicationmodel}\neq\emptyset$.
We write $\mscsetofmodel{\acommunicationmodel}$ for the set of all MSCs linearisable in $\acommunicationmodel$.
\end{definition}

In other words, an MSC $M$ belongs to $\mscsetofmodel{\synchmodel}$ if
all send events are immediately followed by their corresponding receive events.

In our setting, global types are automata that describe a language of MSCs. 
An \emph{arrow} is a triple $(p,q,m)\in\Procs\times\Procs\times\Messages$ with $p\ne q$;
we often write $\marrow{p}{q}{m}$ instead of $(p,q,m)$, and write $\labelalphabet$
to denote the finite set of arrows.
\begin{definition}[Global Type]
A global type $\gt$ is a deterministic finite state automaton over the alphabet $\labelalphabet$.
\end{definition}

Informally, weakly-realizable means that it is realizable and deadlock is allowed.
%% definire da linguaggio di G
\begin{definition}[Weakly-realizable]
It is the 
weak closure $L^w$ of a set $L$ of global type contains all the global type $L$ weakly implies.
The set $L$ is weakly realizable iff $L=L^w$. The global type-graph $G$ is said to be weakly
realizable if the set $L(G)$ of global type is. Thus, a weakly realizable graph already contains
all the implied scenarios.
\end{definition}


\begin{definition}[Relaxed Post Correspondence Problem (RPCP) problem]
Given $\{(v_1, w_1), (v_2, w_2), ..., (v_r, w_r)\}$, determinate whether there are indices $i_1, ..., i_m$
such that $x_{i_1}...x_{i_n}=y_{i_1}...y_{i_m}$, where $x_{i_j}, y_{i_j} \in \{v_{i_j},w_{i_j}\}$, for some
index $i_l \  x_{i_l}\neq y_{i_l}$, and for all $j\leq m,y_{i_1}...y_{i_j}$ is a prefix of $x_{i_1}...x_{i_j}$.
The relaxed problem essentially says that must have the left side of the equivalence that grows more than the right side,
we must have also some tuple that are different, and that we can choose whatever side of tuple for the composition
of the string.
\end{definition}

RPCP is proven to be an undecidable problem by reduction from the classic
PCP problem by Alur,~et~al.~\cite{alur2005realizability}.

A global type defines a language of MSCs in two different ways, one  
existential and one universal. Let $\labellanguageof{\gt}$ be the set of  
sequences of arrows $w$ accepted by $\gt$. Note that for $w \in \Arrows^*$,  
the function $w \mapsto \labeltomsc{w}$ with  
$\labeltomsc{w} \in \mscsetofmodel{\synchmodel}$ is not injective, as two  
arrows with disjoint pairs of processes commute. We write $w_1 \sim w_2$ if  
$\labeltomsc{w_1} = \labeltomsc{w_2}$, and $[w]$ for the equivalence class  
of $w$ with respect to $\sim$.  

The existential MSC language $\existentialmsclanguageof{\gt}{}$ of a global  
type $\gt$ is the set of MSCs that admit at least one representation as a  
sequence of arrows in $\labellanguageof{\gt}$, and the universal MSC  
language $\universalmsclanguageof{\gt}$ of a global type $\gt$ is the set of  
MSCs whose representations as sequences of arrows are all in  
$\labellanguageof{\gt}$:  

$$
\existentialmsclanguageof{\gt} \eqdef \{\labeltomsc{w} \mid 
w \in \labellanguageof{\gt}\} \qquad
\universalmsclanguageof{\gt} \eqdef \{\labeltomsc{w} \mid 
[w] \subseteq \labellanguageof{\gt}\}.
$$

\begin{definition}[Commutation-closed]
    A global type $\gt$ is \emph{commutation-closed} if
    $$
    \existentialmsclanguageof{\gt} = \universalmsclanguageof{\gt}.
    $$
\end{definition}

We write $\msclanguageof{\gt}{}$ for the common language.


\begin{definition}[$G^*$]\label{def:gstar}
Given a set of MSCs $M$, define $G^*$ to be the set of global types
$\{G\ |\ \forall m \in M.\ \msclanguageof{M}{}\in\msclanguageof{\gt}{}\}$.
\end{definition}

%% scrivere proprio gli stati e le transizioni in maniera formale,
% fare altra def per parti intermedie
\begin{definition}[The $L^*$ global type]\label{def:lstar}
Assume to have a finite set $L$ of MSCs, where  
$L = \{m\ |\ m \in M_{\text{synch}}\}$. Let $G^*$ defined as
Definition~\ref{def:gstar}.
We define the global type $L^*_{N}$ as the following automata
$\mathcal A = (Q,\Sigma, \delta, l_0, F)$ where:
\begin{itemize}
  \item $Q$ is a set of states where there are $\{v_I,v_T\}$ united with
  all the state of the global types in $G^*$;
  \item $\Sigma$ is an alphabet composed by the $\epsilon$ united with
  all the labels of the global types in $G^*$;
  \item $\delta:Q\times\Sigma\to 2^Q$ is the transition relation, and
  we correlate: 
  \begin{itemize}
    \item $v_I$ to all initial states of the global types in $G^*$;
    \item each final state of every global type in $G^*$ to $v_T$;
    \item for all $G, G' \in G^*$, connect each final state of $G$
    to the initial state of $G'$.
  \end{itemize}
  \item $l_0 = v_I$ is the initial state;
  \item and F is the set accepting state compose by only $v_F$.
\end{itemize}
Finally, $L^*$ is obtained by the determinisation of $L^*_{N}$.
\end{definition}
Informally, we define $L^*$ as the global type that represent all the
possible MSCs execution of a set of MSCs.

\subsection{Proof}

\begin{lemma}\label{lemma:minsynch}
The MSC $M_i^n$ with $n \in \{0,1\}$ is in $\mscsetofmodel{\synchmodel}$.
\end{lemma}

\begin{proof}
By Definition~\ref{def:linearisable-msc} and Definition~\ref{def:synch},
we just need to show a linearization with all the send operations
followed by the corresponding receive operations, that is
$$
\{~!m_1?m_1!m_2?m_2!m_3?m_3!m_4?m_4!m_5?m_5!m_6?m_6...~\}.
$$
\end{proof}

\begin{theorem}
Given a global type $G$, checking if $G$ is weakly-realizable is undecidable.
\end{theorem}

\begin{proof}
The proof proceeds via a reduction from the RPCP problem. Let's define some useful
elements for the proof.

Given an instance $\Delta = \{(v_1, w_1), \ldots, (v_m, w_m)\}$ of RPCP, we  
construct a set $L$ of MSCs over four processes as follows. For each pair  
$(v_i, w_i)$, we define two MSCs, $M^0_i$ and $M^1_i$, as illustrated in  
Fig.~\ref{fig:msc}.  
In $M^0_i$, process~1 synchronously sends the label $(i, 0)$ to process~2,  
then transmits the index $i$ to process~4. Subsequently, process~4 sends  
$(i, 0)$ synchronously to process~3. After these control messages, process~2  
sends the characters of $v_i$ synchronously to process~3 (where $c$ is the  
length of $v_i$, and $d$ that of $w_i$, as indicated in the figure).  
The MSC $M^1_i$ is constructed analogously, replacing $v_i$ with $w_i$ and
0 with 1.  
Observe that the communication graph of each MSC is strongly connected and  
involves all four processes. Therefore, the MSC graph $L^*$ derived from $L$  
is bounded. With the set $L$, and following the Definition~\ref{def:lstar},
we can construct the global type $L^*$.

% \begin{figure}[h]
% \centering
% \begin{tikzpicture}[->, node distance=24mm, on grid, auto]
%   \node[state] (q0) {$q_0$};
%   \node[state] (q1) [right=of q0] {$q_1$};
%   \node[state] (q2) [right=of q1] {$q_2$};
%   \node[state] (q3) [right=of q2] {$q_3$};
%   \node[state] (q4) [right=of q3] {$q_4$};
%   \node[state] (q5) [right=of q4] {$q_x$};
%   \node[state] (q6) [right=of q5] {$q_y$};

%   \path (q0) edge[] node[above] {\arrmess{p}{q}{(i,n)}} (q1);
%   \path (q1) edge[] node[above] {\arrmess{p}{s}{i}} (q2);
%   \path (q2) edge[] node[above] {\arrmess{s}{r}{(i,n)}} (q3);
%   \path (q3) edge[] node[above] {\arrmess{q}{r}{x_i^1}} (q4);
%   \path (q4) edge[] node[above] {\arrmess{q}{r}{...}} (q5);
%   \path (q5) edge[] node[above] {\arrmess{q}{r}{x_i^c}} (q6);
% \end{tikzpicture}
% \caption{The global type $G^n_i$ where $n\in\{0,1\}$ and $x\in\{v,w\}$.}
% \label{fig:gtype}
% \end{figure}

We need to prove:
\begin{center}
$\Delta \in \text{RPCP}$ if and only if the global type $L^*$ is not weakly-realizable.
\end{center}

\begin{itemize}

  \item[$\Rightarrow$] Let's assume that
  $R = (i_1, a_1, b_1, i_2, a_2, b_2, \ldots, i_m, a_m, b_m)$ are the indices
  for a solution to a generic RPCP problem instance, and the bits $a_j$ and 
  $b_j$ indicate which string ($v_{i_j}$ or $w_{i_j}$) is chosen to go into 
  the two (left and right)
  long strings. Consider the new MSCs $M$ and $M'$ obtained from the sequences
  $M = M^{a_1}_{i_1} \cdots M^{a_m}_{i_m}$ and $M' = M^{b_1}_{i_1} \cdots M^{b_m}_{i_m}$.
  Executions of both of these (sequences of) MSCs must exist in any
  realization of $L^*$. We then look at the projections $M|_1$, $M|_2$, $M|_3$,
  and $M|_4$ of $M$, and $M'|_1$, $M'|_2$, $M'|_3$, $M'|_4$ of $M'$ onto the
  4 processes. Now consider an MSC $M''$ formed from $M'|_1$, $M'|_2$,
  $M|_3$, and $M|_4$. The claim is that the combined MSC $M''$ is weakly
  implied by $L^*$. By definition, the only thing to establish is that $M''$
  is indeed an MSC, in the sense that it is acyclic, well-formed, complete
  and synchronous. This and the previous MSCs are all in
  $\mscsetofmodel{\synchmodel}$ thanks to Lemma~\ref{lemma:minsynch}.

  The only new situation in terms of communication in $M''$ is the
  communication between $P_1$ and $P_4$, and between $P_2$ and $P_3$.
  But the communication between $P_1$ and $P_4$ is consistent in
  $M'|_1$ and $M|_4$ (i.e., the sequence of messages sent from $P_1$ to
  $P_4$ in $M'|_1$ is equal to the sequence of messages received in $M|_4$),
  and the communication between $P_2$ and $P_3$ is consistent in
  $M'|_2$ and $M|_3$ because $R$ is a solution to the RPCP.

  Furthermore, the acyclicity of $M''$ follows from the property of the
  solution that the string formed by the first $j$ words on processes 1
  and 2 is always a prefix of the string formed by the first $j$ words
  on processes 3 and 4. Consequently, each message from $P_1$ to $P_4$
  is sent before it needs to be received.

  But note that $M''$ cannot itself be in $L^*$ because there must be
  some index $i_j$ where $a_j \neq b_j$, and no MSC exists in $L$ where,
  after process 1 announces the index, what process 2 sends is not
  identical to what process 3 receives.

  
  \item[$\Leftarrow$] If the language $L^*$ is not realizable,...

\end{itemize}

\textbf{Lemma.} Each $\mathcal{L}_i^n$ is a synch-language.

\begin{proof}
We show that the MSCs in $\mathcal{L}_i^n$ admit only synchronous linearizations.
Due to the presence of ACK messages, send and receive
events are closely paired. The only possible reordering involves
the ACK from \arrmess{p}{s}{i}, ...
\end{proof}


\end{proof}


