\documentclass{article}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{newlfont}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{lipsum}
\usepackage{tikz}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{xcolor, colortbl}
\usepackage{pgfgantt}
\usepackage{graphicx}
\usepackage{dirtree}
\usepackage{todonotes}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{msc}
\usepackage{subcaption}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\usetikzlibrary{automata,arrows,positioning}
\pgfmathtruncatemacro\distance{1}

\definecolor{green}{rgb}{0,0.5,0}
\definecolor{red}{rgb}{1,0,0}
\definecolor{yellow}{rgb}{0.5,0.5,0}
\definecolor{codeashgrey}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codeashgrey},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

\theoremstyle{definition}
\newtheorem{exmp}{Example}[section]

\title{Final report: On the Implementability of Global Types}
\author{Student: Gabriele Genovese\\Supervisor: Cinzia Di Giusto}
\date{\today}

\begin{document}

\maketitle

\section*{Abstract}
Behavioral types help define how information is exchanged in distributed systems.  
Multiparty Session Types (MPST) describe interactions between multiple participants  
using global protocols and their local versions. Ensuring correct implementation,  
including deadlock freedom and session conformance, is a key concern in MPST.  
Most research focuses on peer-to-peer communication, but real-world systems use  
different models like mailbox-based or causally ordered messaging. A major challenge  
is that a protocol working in one model may fail in another.  
In this work, we explore a flexible MPST framework that adapts to various communication models,
including unordered, causally ordered, and bus-based messaging. Implementability  
is defined through quasi-synchronous (QS) local types, which ensure deadlock freedom  
and prevent orphan messages while maintaining compatibility with synchronous semantics.  
This approach extends MPST beyond peer-to-peer models, making it useful for  
quasi-synchronous systems, which offer better guarantees for verifying correctness.

\newpage

\tableofcontents

\newpage

\input{content/introduction}

\input{content/preliminary}

\input{content/sota}

\input{content/proof}

\newpage

\bibliographystyle{plain}
\bibliography{ref}

\end{document}