\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

First and foremost, I would like to thank the researchers who actively 
collaborated in writing this thesis. Above all, I wish to thank my 
internship supervisor at the laboratory in Nice, Prof. Cinzia Di Giusto, 
who, with great perseverance, demanded formal accuracy in every detail. 
Thanks to her meticulousness, I learned a lot about this profession and 
understood how important it is never to overlook anything.

I would also like to thank Prof. Ivan Lanese for his support during the 
writing of this thesis and for his valuable corrections. I am also grateful 
to Prof. Etienne Losez for the helpful and detailed preliminary discussions 
about the contributions presented in this work.

I wish to thank my parents, Riccardo and Paola, for always being there for me, 
for their unconditional love, and for giving me the freedom to make mistakes 
and grow.

I thank Eleonora for her courage in placing trust in me and in our shared life 
project, which we are building with effort but which I believe will become 
the most beautiful thing in my life.

I thank Samuele, who in his own way inspires me through his choices and 
shares with me the hassle of managing our home server.

I thank all those I met during this master’s degree, especially the friends 
from my bachelor’s years and the new ones adopted along the way: Simo Borto, 
Criss, Fede, Andreea, Simo R., Dan, DavideDarioCiro, Santo, Fusillo, Chapo, 
LP, Apo, Mario, Greg, Gaia, Andrea Ianno, and Luca Bassi. I’ve made many 
memories I will carry in my heart, from every exam we studied for together 
to the struggle of finding our own place, which we then called the 
\emph{bunker}.

I am grateful to those who gave me the opportunity to spend a year abroad in 
Nice through the Erasmus program. It was an incredible year that allowed me 
to leave home for the first time and live somewhere other than Bologna. It 
helped me grow and gave me the confidence to say that I can make it on my own. 
Thank you to Erik for sharing this adventure from the start and for being my 
personal trainer. Thanks to all the old and new \emph{ubinetters}. Through 
this experience, I met so many amazing friends to whom I owe a lot. Among the 
first, the group \emph{Metà Stivale} with Silvia, Arianna, Giuse, Niky, 
Sabri, Lucry, and Amira. The family-like atmosphere that was created was 
magical and unrepeatable.

During the wild summer period, I met some of the most fun-loving people I 
know, and I thank you all for the crazy nights, trips, and travels. Thanks 
to the \emph{Carras team}, Mary and Ana, whose flat became a second home. Thanks 
to Bence, Coleen, Dory, Hannah, Marialena, Mesi, Myriam, Pietro, Richi, 
Hugo, Joy, Isaac, Lea, and Pia for sharing this experience with me. I really 
hope to meet each of you again in the future.

Finally, thank you to my lifelong friends, who remain as anchors I can always 
rely on. Special thanks to those from my bachelor’s days: Luca, Fabio, Gian, 
and Volpe. Thanks to my friends from Bologna: Alf, Annac, Annam, Giulietta, 
Vanni, Leti, Marto, and Tommy. Thanks to my scout community, 
\emph{Bologna 13}, to the children of the \emph{branco Waingunga}, 
and to all the scout chief with 
whom I shared that journey. I sincerely hope to return to scouting in the 
future. I feel truly fortunate to have each of these people in my life.
