
% !TEX root =  ../mainPPDP.tex
%!TEX spellcheck = en_GB

In this section, we show that the realisability problem for complementable global types,
given with an explicit complement, is decidable in both the synchronous and the p2p model.
We first show the result for the synchronous model, which is simpler,
and then extend it to the p2p model using a characterisation of implementability in the p2p model.

\subsection{Realisability in the synchronous model}

\begin{theorem}
    \label{thm:decidability-of-implementability-in-synch}
    Given a global type $\gt$ and its complement $\gt'$, it is decidable in PSPACE complexity whether  $\gt$ is deadlock-free realisable in $\synchmodel$.    
%  The  problem is in PSPACE:
  %  \begin{itemize}
    %    \item Input: a global type $\gt$ and a complement $\gt'$ of $\gt$.
      %  \item Question: is ?
    %\end{itemize}
\end{theorem}

\begin{proof}(Sketch)
    Condition~1 of Definition~\ref{def:realisability}
    is equivalent to $$\msclanguageofcc{\productof{\projectionof{\gt}}} \cap \existentialmsclanguageof{\gt'} = \emptyset,$$
    which reduces to checking the emptiness of $$\languageofnfa{\productof{\projectionof{\gt}}\otimes\gt'},$$
    where $\otimes$ is the standard product of NFAs mimicking intersection. This equivalence holds 
    because $\productof{\projectionof{\gt}}$ is commutation-closed. This condition
    can be checked in non-deterministic logarithmic space in the size of the product NFA,
    which yields a PSPACE algorithm if the exponential size NFA  $\preproductof{\projectionof{\gt}}$
    is lazily constructed (note that it is not necessary to determinise the NFA, as we only need to check emptiness). 
    Assuming that Condition~1 of Definition~\ref{def:realisability} holds,
    the second condition of Definition~\ref{def:realisability}  is 
    equivalent to whether all control states 
    of $\preproductof{\projectionof{\gt}}$ that are reachable from the initial state
    can reach a final state, which again can be checked in PSPACE.
\end{proof}

\subsection{Realisability in the $\ppmodel$ model}

%When moving to $\ppmodel$ the proof is significantly more involved and we need to resort  to  an equivalent  characterisation of deadlock-free realisability in the $\ppmodel$ model.

\input{proofs/thm-main-theorem-implementability.tex}    
    
A system $\system$ such that 
$\msclanguageof{\system}{\ppmodel}\subseteq
\prefixclosureof{\mscsetofmodel{\synchmodel}}$
is called \emph{realisable with synchronous communications} (RSC) 
in~\cite{germerie-phd}. Germerie~\emph{et al.}~\cite{germerie-phd}
showed that the RSC property is decidable in PSPACE and that
whether a RSC system may reach a regular set of configurations is
also in PSPACE. As a consequence, conditions~1 and 2 of
Theorem~\ref{thm:main-theorem-realisability} can be checked in PSPACE.
Assuming that $\projectionof{\gt}$ is RSC, 
$\mscsofcfsms{\projectionof{\acceptcompletion{\gt}}}{\ppmodel}$ can be coded as an effective regular set of
executions, as well as $\prefixclosureof{\mscsofcfsms{\projectionof}{\ppmodel}}$,
and Condition~3 of Theorem~\ref{thm:main-theorem-realisability} reduces to checking the inclusion of
two exponential size NFA, which can be checked in PSPACE using a lazy construction of the NFAs.


% It remains to show that the third condition of
% Theorem~\ref{thm:main-theorem-implementability}
% is decidable in PSPACE. 
% This relies on the following lemma that states that 
% condition~3 of Theorem~\ref{thm:main-theorem-realisability} is equivalent to
% being dead-lock free in the sense of Definition~\ref{def:orphan-deadlock-free}.

% \begin{restatable}{lemma}{eventualterminationonmsc}\label{lem:eventual-termination-can-be-checked-on-mscs-for-p2p}
%     $\executionsof{\projectionof{\acceptcompletion{\gt}}}{\ppmodel}=\prefixclosureof{\executionsof{\projectionof{\gt}}{\ppmodel}}$ is equivalent to
%     is equivalent to $\gt$ being deadlock-free in $\ppmodel$.
% \end{restatable}
% \input{proofs/lem-eventual-termination-can-be-checked-on-mscs-for-p2p.tex}

From these results we can conclude that it can be decided in PSPACE that a given $\gt$,
together with a complement hint, is deadlock-free realisable in the $\ppmodel$ model.
 \begin{theorem}\label{thm:decidability-of-implementability-in-p2p}
   
   Given a complementable global type $\gt$ and its complement $\gt'$,     it is decidable in PSPACE complexity whether $\gt$ is deadlock-free realisable in $\ppmodel$.
   
%       The following problem is in PSPACE:
 %\begin{itemize}
  %      \item Input: a complementable global type $\gt$ and a complement $\gt'$ of $\gt$.
    %    \item Question: is $\gt$ deadlock-free realisable in $\ppmodel$?
    %\end{itemize}
\end{theorem}

