\documentclass[11pt,a4paper]{letter}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{parskip}
\geometry{margin=1in}

\begin{document}

\begin{letter}{DisCoTec 2025 Organizing Committee\\
University of Lille\\
Lille, France}

\opening{Dear DisCoTec 2025 Organizing Committee,}

I am writing to apply for a student travel grant to attend DisCoTec 2025, which will be 
held in Lille, France. I am currently pursuing a Master's degree in Computer Science 
at the University of Bologna, and I am also an intern at the I3S Laboratory in Sophia 
Antipolis under the supervision of Prof Cinzia Di Giusto.

DisCoTec represents an exciting opportunity for me as I am planning to pursue a PhD in 
Computer Science theory. Participating in this conference would allow me to meet and exchange 
ideas with established researchers and fellow students, and to immerse myself in 
the research and development of distributed system theory. I believe 
this exposure will be highly beneficial as I refine my research interests and look 
toward my academic future.

I am particularly interested in the tutorials scheduled at the conference. In 
particular, ``A Choreographic View of Smart Contracts'' by Emilio Tuosto aligns 
perfectly with my academic focus and my interest in formal methods and distributed 
systems. Similarly, I really look forward to the ``Tutorial on EnOSlib'' by Baptiste Jonglez 
as it seems it promises valuable 
insight into reproducible experimentation and system evaluation, which I find very interesting for 
both practice and theory.

Being able to attend DisCoTec 2025 would be an important step in my academic journey,
as it would be my first participation on a conference. 
It would not only strengthen my background in core research areas but also help me 
build relationships with the international research community, which I think is very
important for my future as a PhD student.

Thank you for considering my application. I hope to contribute actively and 
enthusiastically to the event.

\closing{Sincerely,\\[1.5ex]Gabriele Genovese}

\end{letter}

\end{document}
