\documentclass{beamer}
\usetheme{boxes}

\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{listings}

\title{Choreographies for Program Understanding}
\author{Gabriele Genovese et al.}
\date{\today}

\begin{document}

\frame{\titlepage}

% \begin{frame}{Overview}
% \tableofcontents
% \end{frame}


\begin{frame}{Motivation}
\begin{itemize}
  \item Distributed systems use message-passing for coordination.
  \item Understanding communication behavior is challenging.
  \item Existing choreography-based approaches focus on top-down development.
  \item This paper promotes bottom-up extraction of choreographies for program understanding.
\end{itemize}
\end{frame}

\begin{frame}{Goals and Vision}
\begin{itemize}
  \item Automatically extract global views from existing code.
  \item Help understand interaction patterns and detect bugs.
  \item Support debugging through graphical, analyzable global views.
\end{itemize}
\end{frame}

\begin{frame}{Requirements}
\begin{itemize}
  \item Push-button extraction from real code (e.g., Erlang, Go).
  \item Capture both correct and faulty behaviors.
  \item Highlight misbehaviors like deadlocks and races.
  \item Provide readable and precise graphical representations.
\end{itemize}
\end{frame}

\begin{frame}{Challenges}
\begin{itemize}
  \item \textbf{Undecidability:} Complete extraction is impossible in general.
  \item \textbf{Scalability:} Large programs lead to complex choreographies.
  \item \textbf{Approximation:} Over- and under-approximations may introduce noise.
\end{itemize}
\end{frame}

\begin{frame}{Examples}
\begin{itemize}
  \item \textbf{Dining Philosophers:} Detects potential deadlock.
  \item \textbf{Bank Account:} Reveals potential race in balance updates.
\end{itemize}
\vspace{0.5em}
Both examples are processed with the prototype tool \texttt{Chorer}.
\end{frame}

\begin{frame}{Tool Support: Chorer}
\begin{itemize}
  \item Parses Erlang programs and extracts local/global automata.
  \item Uses static analysis and Graphviz DOT for output.
  \item Detects deadlocks; supports heuristics to manage complexity.
\end{itemize}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}
  \item Choreographies can aid in understanding real-world message-passing code.
  \item Bottom-up extraction complements traditional top-down design.
  \item Future work includes better support for loops, recursion, and dynamic actors.
\end{itemize}
\end{frame}

\end{document}
